Automatic Amortized Analysis for Imperative Code

### 2014-03-06 main.tex
fix bug in semantics
more tolls

\Rule{Skip}
{}
{ \Gamma; Q \vdash \code{skip} \dashv Q; \Gamma }

\and \Rule{Assert}
{}
{ \Gamma; Q \vdash \code{assert}(\Delta) \dashv Q; \Gamma, \Delta }

\\ \Rule{AssignP}
{ \Gamma \models \Gamma'[x/x+y] }
{\Gamma; Q + y\cdot q_x \vdash x \gets x + y \dashv Q; \Gamma'}

\and\Rule{AssignM}
{  \Gamma \models \Gamma'[x/x-y]
\\ \Gamma \models c \le y \le x
}
{ \Gamma; Q \vdash x \gets x - y \dashv c \cdot q_x + V; \Gamma' }

\\ \Rule{While}
{ \Gamma, 0 < x ; Q - M_w \vdash S \dashv Q; \Gamma }
{ \Gamma; Q \vdash \code{while}(0 < x)~S \dashv Q; \Gamma, x \le 0 }

\and\Rule{Seq}
{  \Gamma; Q \vdash S_1 \dashv Q'; \Gamma'
\\ \Gamma'; Q' \vdash S_2 \dashv Q''; \Gamma''
}
{ \Gamma; Q \vdash S_1;S_2 \dashv Q''; \Gamma'' }

@@

\Rule{AssignOpK}
{  \Gamma \models \Gamma'[x/x \pm y]
\\ \Gamma \models |y| = k
\\ \mathcal U = \left\{ u \mid \Gamma \models x \pm y \in \inter x u \right\}
\\ \textstyle q'_y = q_y
      + \sum_{u \in \mathcal U} p_{xu}
      - \sum_{v \not\in \mathcal U} q_{xv}
\\ q_{xu} = p_{xu} + r_{xu}
}
{ \textstyle
  \Gamma; Q
  \vdash x \gets x \pm y \dashv
  \sum_u k \cdot r_{xu} + Q'; \Gamma'
  % q_x is not omitted, v_{x} is in fact a shorthand for
  % v_{x0} where 0 is a variable equal to 0.
}

\and \Rule{AssignOp}
{  \Gamma \models \Gamma'[x/x \pm y]
\\ \mathcal U = \left\{ u \mid \Gamma \models x \pm y \in \inter x u \right\}
\\ \textstyle q'_y = q_y
      + \sum_{u \in \mathcal U} q_{xu}
      - \sum_{v \not\in \mathcal U} q_{xv}
}
{ \textstyle
  \Gamma; Q \vdash x \gets x \pm y \dashv Q'; \Gamma'
}

@@

This big step semantics counts the number of iterations of
while loops in a program as it executes it.  Non-deterministically,
the {\sc S:Bubble} rule can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize an
immutable heap with the notation $\circ H$, this kind of heap just
bubbles up through the semantics. In the rules, $X$ stands for
either $H$ or $\circ H$ for some $H$.

@@

\Rule{S:Bubble}
{}
{ (H, S) \Downarrow_0 \circ H }

\and \Rule{S:Skip}
{}
{ (H, \code{skip}) \Downarrow_0 H }

\and \Rule{S:Assign}
{}
{ (H, x \gets e) \Downarrow_0 H[x \gets \llbracket e \rrbracket_H] }

\and \Rule{S:Assert}
{ \Gamma \models H }
{ (H, \code{assert}(\Gamma)) \Downarrow_0 H }

\and \Rule{S:WhileStop}
{ \neg (k < \llbracket x - y \rrbracket_H) }
{ (H, \code{while}(k < x - y)~S) \Downarrow_0 H }

\and \Rule{S:WhileStep}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S; \code{while}(k < x - y)~S) \Downarrow_n X
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{n+1} X }

\and \Rule{S:If}
{  (H, S_i) \Downarrow_n X
\\ i \in \{1, 2\}
}
{ (H, \code{if}(?)~S_1~\code{else}~S_2) \Downarrow_n X }

\and \Rule{S:Seq}
{  (H, S_1) \Downarrow_{n_1} H'
\\ (H', S_2) \Downarrow_{n_2} H''
}
{ (H, S_1; S_2) \Downarrow_{n_1 + n_2} H'' }

\and \Rule{S:SeqBubble}
{ (H, S_1) \Downarrow_n \circ H' }
{ (H, S_1; S_2) \Downarrow_n \circ H' }

>>>>

\Rule{Skip}
{}
{ \Gamma; Q + M_s \vdash \code{skip} \dashv Q; \Gamma }

\and \Rule{Assert}
{}
{ \Gamma; Q + M_a \vdash \code{assert}(\Delta) \dashv Q; \Gamma, \Delta }

% \\ \Rule{AssignP}
% { \Gamma \models \Gamma'[x/x+y] }
% {\Gamma; Q + M_s + y\cdot q_x \vdash x \gets x + y \dashv Q; \Gamma'}

% \and\Rule{AssignM}
% {  \Gamma \models \Gamma'[x/x-y]
% \\ \Gamma \models c \le y \le x
% }
% { \Gamma; Q + M_s \vdash x \gets x - y \dashv c \cdot q_x + V; \Gamma' }

\\ \Rule{While}
{ \Gamma, 0 < x ; Q - M_w^2 \vdash S \dashv Q; \Gamma }
{ \Gamma; Q + M_w^1 \vdash \code{while}(0 < x)~S \dashv Q - M_w^3; \Gamma, x \le 0 }

\and\Rule{Seq}
{  \Gamma; Q - M_s^1 \vdash S_1 \dashv Q'; \Gamma'
\\ \Gamma'; Q' - M_s^2 \vdash S_2 \dashv Q''; \Gamma''
}
{ \Gamma; Q \vdash S_1;S_2 \dashv Q''; \Gamma'' }

@@

\Rule{AssignOpK}
{  \Gamma \models \Gamma'[x/x \pm y]
\\ \Gamma \models |y| = k
\\ \mathcal U = \left\{ u \mid \Gamma \models x \pm y \in \inter x u \right\}
\\ \textstyle q'_y = q_y
      + \sum_{u \in \mathcal U} p_{xu}
      - \sum_{v \not\in \mathcal U} q_{xv}
\\ q_{xu} = p_{xu} + r_{xu}
}
{ \textstyle
  \Gamma; Q + M_s
  \vdash x \gets x \pm y \dashv
  \sum_u k \cdot r_{xu} + Q'; \Gamma'
  % q_x is not omitted, v_{x} is in fact a shorthand for
  % v_{x0} where 0 is a variable equal to 0.
}

\and \Rule{AssignOp}
{  \Gamma \models \Gamma'[x/x \pm y]
\\ \mathcal U = \left\{ u \mid \Gamma \models x \pm y \in \inter x u \right\}
\\ \textstyle q'_y = q_y
      + \sum_{u \in \mathcal U} q_{xu}
      - \sum_{v \not\in \mathcal U} q_{xv}
}
{ \textstyle
  \Gamma; Q + M_s \vdash x \gets x \pm y \dashv Q'; \Gamma'
}

@@

This big step semantics counts the number of iterations of
while loops in a program as it executes it.  Non-deterministically,
the {\sc S:Abort} rule can be called to interrupt a computation.
It has the same effect as raising an exception.  We symbolize a
frozen heap with the notation $\circ H$, this kind of heap just
bubbles up through the semantics. In the rules, $X$ stands for
either $\circ$ or $H$.

@@

\Rule{S:Abort}
{ }
{ (H, S) \Downarrow_0 \circ }

\and \Rule{S:Skip}
{ }
{ (H, \code{skip}) \Downarrow_{M_s} H }

\and \Rule{S:Assign}
{ }
{ (H, x \gets e) \Downarrow_{M_s} H[x \mapsto \llbracket e \rrbracket_H] }

\and \Rule{S:Assert}
{ H \models \Gamma }
{ (H, \code{assert}(\Gamma)) \Downarrow_{M_a} H }

\and \Rule{S:WhileStop}
{ \neg (k < \llbracket x - y \rrbracket_H) }
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^3} H }

\and \Rule{S:WhileStep}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{q_1} H'
\\ (H', \code{while}(k < x - y)~S) \Downarrow_{q_2} X
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot q_1 \cdot M_w^2 \cdot q_2} X }

\and \Rule{S:WhileStepAbort}
{  k < \llbracket x - y \rrbracket_H
\\ (H, S) \Downarrow_{q_1} \circ
}
{ (H, \code{while}(k < x - y)~S) \Downarrow_{M_w^1 \cdot q_1 } \circ }

\and \Rule{S:If}
{  (H, S_i) \Downarrow_q X
\\ i \in \{1, 2\}
}
{ (H, \code{if}(?)~S_1~\code{else}~S_2) \Downarrow_{M_c^i \cdot q} X }

\and \Rule{S:Seq}
{  (H, S_1) \Downarrow_{q_1} H'
\\ (H', S_2) \Downarrow_{q_2} X
}
{ (H, S_1; S_2) \Downarrow_{M_s^1 \cdot q_1 \cdot M_s^2 \cdot q_2} X }

\and \Rule{S:SeqAbort}
{ (H, S_1) \Downarrow_q \circ }
{ (H, S_1; S_2) \Downarrow_q \circ }
